Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-15 writedown/period.noneq.msm/paper.tex
make the notations more consistent. backup before major structure change

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{red} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
% \newcommand{\fwd}[0]{\textrm{fw}}
% \newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\fwd}[0]{+}
\newcommand{\bwd}[0]{-}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_n\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, n\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{n\times n}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^n L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^T(t)\cdot \vect p(t)
\end{align}
where $\vect L(t)^T$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $n$-vector denoting the probability distribution on $S$ at time $t$, and
$p_i(t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{n\times n}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^T(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^n \Phi_{ij}(t)  =  1$,
that is, $\myphi^T(t)$ is a
stochastic matrix satisfying $\myphi^T(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^T\in \mathbb R^n$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L(t+T)^T \vect Y(t)=\vect L(t)^T\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.
Since $\myphi(T)$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with left eigenvector $\vect e=(1,\ldots,1)^T$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty,
\end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,n$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix 
\begin{equation}\label{P}
\vect P=\myphi^T(T)
\end{equation}
 is discrete in time but in space it still is a fine-scale description of the transport properties of the
dynamics with a very large number $n$ of states.  Now we want to coarsen our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $k\ll n$
\emph{macrostates} that should be the metastable states of the system: The resulting $k\times k$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that given the molecular system under consideration exhibits metastable conformations then it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescale significantly shorter than the metastable timescales.
Then if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $k$ core sets $C_1,\ldots, C_k\subset S$ that are appropriate metastable sets. 
Following \cite{A19-31,schuette2011markov} the process $(X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $k$ states associated with the sets $C_j$, $j=1,\ldots,k$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(X_m)$, i.e., trajectories of $(X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\[
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j|X_m=i),
\] 
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
From the last equation we get that the stationary distribution of the milestone process is given by
\[
\hat{\mu}(j)=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\]
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}(j)$.

Following \cite{discreteTPT} one also has to consider the forward committor $q^\fwd_j(i)$ identical to the probability  that conditioned on $X_m=i$ the next core set to be hit will be $C_j$.
The forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$ can be computed from $P$ by solving the linear equations \cite{discreteTPT}
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\label{qfwd}\\
q^\fwd_j(i) & =  1,\quad i\in C_j\nonumber\\
q^\fwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $C=S\setminus\cup_j C_j$, and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\label{qbwd}\\
q^\bwd_j(i) & =  1,\quad i\in C_j\nonumber \\
q^\bwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu_i P_{ij}/\mu_j$.

\recheck{
We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jl}=\prob_\mu\Big(\hat{X}_{m+1}=l\vert \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ is associated with the milestone process $(\hat{X}_m)$ in terms of committors by}
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jl}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_l \rangle_\mu,\qquad j\not= l, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{l\not=j} \hat{\vect P}_{jl}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u_i v_i \mu_i$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}(l,0)$ of being in state $l$ at time $t=0$ then its evolution $\hat{p}(l,t)$ in time is discrete in multiples of period $T$ and given by
\recheck{
\begin{equation}\label{eq:num-29}
\hat{p}(l,T)=\sum_j \hat{p}(j,0)\,\,\hat{\vect P}_{jl}.
\end{equation}}
In our approach $p(j,mT)$ is a good approximation of $\prob(\hat{X}_m=j)$.

\textbf{Remark 1:} Our definition of the milestoning process in terms of the process $(X_m)$ generated by $P$ guarantees that we can directly compute $\hat{P}$ via (\ref{hatP_prob})  from trajectories of $(X_m)$ without computing the committor functions. 
This is of importance if the spatial discretization underlying $(X_m)$ is fine enough, because then the kinetics of $(X_m)$
approximates the original kinetics of $(x_{mT})$ so that we can directly compute   $\hat{P}$ via (\ref{hatP_prob})  from NEMD trajectories without computing $P$ first (which substantially simplified the MSM building if the core sets are already known).

\textbf{Remark 2:}
Following \cite{A19-31} we can also define another pair of stochastic MSM matrices: 
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)},
\end{align}
that are connected to $\hat{P}$ by the following identity
\[
\hat{P}=\hat{T}-\hat{M}+\id,
\]
That can be seen by means of direct computation: For the off-diagonal entries we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{ij} - \hat M_{ij}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \hat P_{jk}, \quad j\neq k
\end{align}
Stochasticity yields
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$ for the diagonal entries.  Furthermore, as shown in the Appendix~\ref{sec:app-prove}, $\hat{T}$ as well as $\hat{M}$ can be computed from trajectories without need to have the committors. 
The importance of the pair $\hat{T}$ and $\hat{M}$ for MSM building comes from the following observation: The main NEMD relaxation timescales are given by the dominant eigenvalues of $P$ \cite{A19-31, A19-1}.
These dominant eigenvalues can be approximated by discretizing the related eigenvalue problem $Pu=\lambda u$ by means of a Galerkin approximation with the finite dimensional ansatz space 
spanned by the forward committors $q^+_j$, $j=1,\ldots,k$ together  with the finite dimensional test function space spanned by the backward committors $q^-_j$, $j=1,\ldots,k$ (test functions multiplied from the left by the inner product $\langle\cdot,\cdot\rangle_\mu$). The thus discretize eigenproblem takes the form  of a generalized eigenproblem
\begin{align}
  \label{eq:msm-gen-ev}
\hat{T}\hat{u}=\hat{\lambda}\hat{M}\hat{u}.  
\end{align}
For the reversible case it is known that its $k$ eigenvalues $\hat{\lambda}$ are very good approximation of the dominant eigenvalues $\lambda$ of the original problem if the core sets are proper metastable sets~\cite{Eigenvalues}. Whether this is true for the non-reversible case is not known yet, but if the deviation from reversibility is weak and the dominant eigenvalues of $P$ are real-valued then the results should hold analogously, see~\cite{A19-31}, Thm.~4.19.

>>>>

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{red} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
% \newcommand{\fwd}[0]{\textrm{fw}}
% \newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\fwd}[0]{+}
\newcommand{\bwd}[0]{-}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}
\newcommand{\dih}{\textrm{dih}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_N\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, N\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{N\times N}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^N L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^T(t)\cdot \vect p(t)
\end{align}
where $\vect L^T(t)$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $N$-vector denoting the probability distribution on $S$ at time $t$, $p_i(t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{N\times N}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^T(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^N \Phi_{ij}(t)  =  1$,
that is, $\myphi^T(t)$ is a
stochastic matrix satisfying $\myphi^T(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^T\in \mathbb R^N$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

@@

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L^T(t+T) \vect Y(t)=\vect L^T(t)\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.
Since $\myphi(T)$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with left eigenvector $\vect e=(1,\ldots,1)^T$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty,
\end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

@@

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,N$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix 
\begin{equation}\label{P}
\vect P=\myphi^T(T)
\end{equation}
 is discrete in time but in space it still is a fine-scale description of the transport properties of the
dynamics with a very large number $N$ of states.  Now we want to coarsen our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $K\ll N$
\emph{macrostates} that should be the metastable states of the system: The resulting $K\times K$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that given the molecular system under consideration exhibits metastable conformations then it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescale significantly shorter than the metastable timescales.
Then if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov} the process $(\tilde X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\[
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j|X_m=i),
\]
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
\recheck{COMMENT: Here it is a \emph{conditional} probability,
  due to the time-homogeneity,
  it does not matter whether it is defined from quasi-stationary probability
  or not.}
From the last equation we get that the stationary distribution of the milestone process is given by
\[
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\]
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

Following \cite{discreteTPT} one also has to consider the forward committor $q^\fwd_j(i)$ identical to the probability  that conditioned on $\tilde X_m=i$ the next core set to be hit will be $C_j$.
The forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$ can be computed from $P$ by solving the linear equations \cite{discreteTPT}
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\label{qfwd}\\
q^\fwd_j(i) & =  1,\quad i\in C_j\nonumber\\
q^\fwd_j(i) & =  0,\quad i\in C_k,k\not=j\nonumber
\end{align}
where $C=S\setminus\cup_j C_j$, and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\label{qbwd}\\
q^\bwd_j(i) & =  1,\quad i\in C_j\nonumber \\
q^\bwd_j(i) & =  0,\quad i\in C_k,k\not=j\nonumber
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu(i) P_{ij}/\mu(j)$.

\recheck{
We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ is associated with the milestone process $(\hat{X}_m)$ in terms of committors by}
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\recheck{
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$.

\textbf{Remark 1:} Our definition of the milestoning process in terms of the process $(\tilde X_m)$ generated by $P$ guarantees that we can directly compute $\hat{P}$ via (\ref{hatP_prob})  from trajectories of $(\tilde X_m)$ without computing the committor functions. 
This is of importance if the spatial discretization underlying $(\tilde X_m)$ is fine enough, because then the kinetics of $(\tilde X_m)$
approximates the original kinetics of $(x_{mT})$ so that we can directly compute   $\hat{P}$ via (\ref{hatP_prob})  from NEMD trajectories without computing $P$ first (which substantially simplified the MSM building if the core sets are already known).

\textbf{Remark 2:}
Following \cite{A19-31} we can also define another pair of stochastic MSM matrices: 
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j},
\end{align}
that are connected to $\hat{P}$ by the following identity
\[
\hat{P}=\hat{T}-\hat{M}+\id,
\]
That can be seen by means of direct computation: For the off-diagonal entries we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{jk} - \hat M_{jk}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \hat P_{jk}, \quad j\neq k
\end{align}
Stochasticity yields
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$ for the diagonal entries.  Furthermore, as shown in the Appendix~\ref{sec:app-prove}, $\hat{T}$ as well as $\hat{M}$ can be computed from trajectories without need to have the committors. 
The importance of the pair $\hat{T}$ and $\hat{M}$ for MSM building comes from the following observation: The main NEMD relaxation timescales are given by the dominant eigenvalues of $P$ \cite{A19-31, A19-1}.
These dominant eigenvalues can be approximated by discretizing the related eigenvalue problem $Pu=\lambda u$ by means of a Galerkin approximation with the finite dimensional ansatz space 
spanned by the forward committors $q^+_j$, $j=1,\ldots,K$ together  with the finite dimensional test function space spanned by the backward committors $q^-_j$, $j=1,\ldots,K$ (test functions multiplied from the left by the inner product $\langle\cdot,\cdot\rangle_\mu$). The thus discretize eigenproblem takes the form  of a generalized eigenproblem
\begin{align}
  \label{eq:msm-gen-ev}
\hat{T}\hat{u}=\hat{\lambda}\hat{M}\hat{u}.  
\end{align}
For the reversible case it is known that its $k$ eigenvalues $\hat{\lambda}$ are very good approximation of the dominant eigenvalues $\lambda$ of the original problem if the core sets are proper metastable sets~\cite{Eigenvalues}. Whether this is true for the non-reversible case is not known yet, but if the deviation from reversibility is weak and the dominant eigenvalues of $P$ are real-valued then the results should hold analogously, see~\cite{A19-31}, Thm.~4.19.

\textbf{Remark 3:}
The pair $\hat{T}$ and $\hat{M}$, or the matrix $\hat{T}\hat{M}^{-1}$ (if all of its entries are positive such that it is a stochastic matrix), thus can also be taken as MSM transition matrices. In the case of  $\hat{T}$ and $\hat{M}$ the MSM modelling error results from Galerkin discretization, while the MSM modelling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$.
\recheck{As a matter fact, if $\hat X_m$ is Markovian, the identity of the transition matrices holds: $\hat P = \hat T \hat M^{-1}$ (see~Appendix~\ref{sec:app-ptm} for the prove).
Therefore, in practice, the deviation of the eigenvalues of $\hat P$ from those of $\hat{T}\hat{M}^{-1}$ indicates the non-Markovianity implied in the process $\hat{X}_m$.}
The lesson learnt in the reversible case are that $\hat{P}$ is better suited to approximate the molecular kinetics on large timescales while $\hat{T}$ and $\hat{M}$ should be used for approximating eigenvalues and long relaxation timescales.
